# pre-class


Make sure you commit this often with meaningfull messages. 

### Background

The exponential distribution is defined by its cumulative distribution function
\(F(x) = 1-e^{-\lambda x}\)

The R function ***rexp()*** generates random variables with an exponential distribution. For example 
<center><strong>rexp(n=10, rate=5)</strong> </center>

results in 10 exponentially distributed numbers with a rate \(\lambda=5\). If you leave out the 5 and just have
<center><strong>rexp(n=10) </strong></center>
then this results in 10 exponentially distributed numbers with a rate \(\lambda=1\), this is also referred to as the "standard exponential distribution". 

### Part 1


1. Generate 200 random values from the standard exponential distribution and store them in a vector `exp.draws.1`.  Find the mean and standard deviation of `exp.draws.1`.
```{r}
#This vector holds 200 random values from the standard exponential distribution.
exp.draws.1 <- rexp(n = 200)
#This is the calculation of the mean and standard deviation of the just created vector.
mean(exp.draws.1)
sd(exp.draws.1)
```
2. Repeat, but change the rate to 0.2, 5, 7.3 and 10, storing the results in vectors called  `exp.draws.0.2`,  `exp.draws.5`,  `exp.draws.7.3` and  `exp.draws.10`.
```{r}
#The following vectors hold 200 random values from the exponential distribution with varying rate, with their means and standard deviations calculated.

#With rate = 0.2
exp.draws.0.2 <- rexp(n = 200, rate = 0.2)
mean(exp.draws.0.2)
sd(exp.draws.0.2)

#With rate = 5
exp.draws.5 <- rexp(n = 200, rate = 5)
mean(exp.draws.5)
sd(exp.draws.5)

#With rate 7.3
exp.draws.7.3 <- rexp(n = 200, rate = 7.3)
mean(exp.draws.7.3)
sd(exp.draws.7.3)

#With rate 10
exp.draws.10 <- rexp(n = 200, rate = 10)
mean(exp.draws.10)
sd(exp.draws.10)
```

3. The function `plot()` is the generic function in R for the visual display of data. `hist()` is a function that takes in and bins data as a side effect. To use this function, we must first specify what we'd like to plot.
    a. Use the `hist()` function to produce a histogram of your standard exponential distribution.
```{r}
#Histogram of results from the generated random values from the standard exponential distribution.
hist(exp.draws.1)
```
<br />
The bins each have size 0.5, and the largest bin is the one from 0 to 0.5. The size of the bins then tend to get smaller and smaller, though there are some exceptions.
<br>
<br>
    b. Use `plot()` with this vector to display the random values from your standard distribution in order.
```{r}
#Plot of the generated random values from the standard exponential distribution.
plot(exp.draws.1)
```
<br />
In this plot, the index of the simulated value is plotted against its value. Most of the values are less than 1, as you would expect from the histogram plot.
<br>
    c. Now, use `plot()` with two arguments -- any two of your other stored random value vectors -- to create a scatterplot of the two vectors against each other.
```{r}
plot(exp.draws.1, exp.draws.10)
```
<br>
For any index i, this graphic plots the ith simulated value of the standard exponential distribution against the ith simulated value of the exponential distribution with rate equal to 10.

4. We'd now like to compare the properties of each of our vectors. Begin by creating a vector of the means of each of our five distributions in the order we created them and saving this to a variable name of your choice. Using this and other similar vectors, create the following scatterplots and explain in words what is going on:
    a. The five means versus the five rates used to generate the distribution.
```{r}
#This vector contains the values for the rates used to generate the random values.
rates <- c(1, 0.2, 5, 7.3, 10)
#This vector contains the means of the five sets of simulated values.
means <- c(mean(exp.draws.1), mean(exp.draws.0.2), mean(exp.draws.5), mean(exp.draws.7.3), mean(exp.draws.10))
#This vector contains the standard deviations of the five sets of simulated values.
sds <- c(sd(exp.draws.1), sd(exp.draws.0.2), sd(exp.draws.5), sd(exp.draws.7.3), sd(exp.draws.10))
#This is a plot of the five means versus the five rates used to generate the values.
plot(means, rates)
```
This is a plot of the means versus the rates, generally the mean is close to equal to 1 divided by the rate.
    b. The standard deviations versus the rates.
```{r}
#This is a plot of the standard deviations versus the rates.
plot(sds, rates)
```
    c. The means versus the standard deviations.
```{r}
#This is a plot of the means versus the standard deviations.
plot(means, sds)
```

For each plot, explain in words what's going on.

### Part II (PHP 2560 Only)


5. R's capacity for data and computation is large to what was available 10 years ago. 
    a. To show this, generate 1.1 million numbers from the standard exponential distribution and store them in a vector called `big.exp.draws.1`. Calculate the mean and standard deviation.
    b. Plot a histogram of `big.exp.draws.1`.  Does it match the function \(1-e^{-x}\)?  Should it? 
    c. Find the mean of all of the entries in `big.exp.draws.1` which are strictly greater than 1. You may need to first create a new vector to identify which elements satisfy this.
    d. Create a matrix, `big.exp.draws.1.mat`, containing the the values in 
`big.exp.draws.1`, with 1100 rows and 1000 columns. Use this matrix as the input to the `hist()` function and save the result to a variable of your choice. What happens to your data?
    e. Calculate the mean of the 371st column of `big.exp.draws.1.mat`.
    f. Now, find the means of all 1000 columns of `big.exp.draws.1.mat` simultaneously. Plot the histogram of column means.  Explain why its shape does not match the histogram in problem 5b).
   
