---
title: "Simulations Pre-Class Project"
date: "Due March 13, 2017 at 5:00pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:


With this project we will simulate a famoues probability problem. This will not require knowledge of probability or statistics but only the logic to follow the steps in order to simulate this problem. This is one way to solve problems by using the computer. 

 1. **Gambler's Ruin**: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:
    a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet.
```{r}
n_hands <- function(n, bankroll, bet, odds, bust = TRUE){
  #This function will give you the amount of money you have if you play n hands with bankroll being the amount of money you begin with betting the amount bet each hand. The odds of winning is given by odds (should be between 0 and 1). You also have the option of deciding whether or not you can bust,
  for(i in 1:n){
    win <- rbinom(1,1,odds)
    if(win == 1){
      bankroll = bankroll + bet
    }else{
      bankroll = bankroll - bet
    }
    if(bankroll == 0 & bust == TRUE){
      #If you run out of money, the function returns 0 and the number of the hand where you bankrupted.
      return(c(0,i))
    }
  }
  bankroll
}
#For 5000 simulations, we run the previous function for 100 hands, with 1000 in the bankroll, 100 per bet and 0.5 for the odds since the game is fair. The number of times that you go bust in 100 hands is recorded, and divided by 5000 to give an estimate of the probability that you will bust within 100 hands.
n_bust_100 <- 0
for(i in 1:5000){
  made <- n_hands(100, 1000, 100, 0.5)
  if(length(made) == 2){
    #If length of made is 2, you have busted, by the way the function was defined (it returns 0 and the number of hands until bust).
    n_bust_100 = n_bust_100 + 1
  }
}
prob_bust_100 = n_bust_100/5000
prob_bust_100
``` 
    b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly.
```{r}
#This is similar to the last part, except we limit the bets to 500.
n_bust_500 <- 0
for(i in 1:5000){
  made <- n_hands(500, 1000, 100, 0.5)
  if(length(made) == 2){
    n_bust_500 = n_bust_500 + 1
  }
}
prob_bust_500 = n_bust_500/5000
prob_bust_500
```
    c. the mean time you go bust, given that you go bust within the first 5000 hands.
```{r}
#This time we record the time you go bust in 5000 hands, and then take the mean.
bust_time <- c()
for(i in 1:5000){
  made <- n_hands(5000, 1000, 100, 0.5)
  if(length(made) == 2){
    #Each time you bust, the hand at which you busted is added to the vector bust_time, from which the mean will be calculated.
    bust_time <- c(bust_time, made[2])
  }
}
mean(bust_time)
```
    d. the mean and variance of your bankroll after 100 hands (including busts).
```{r}
#In each simulation of 100 hands, your bankroll is added to a vector. Then the mean and variance is calculated.
end_bankroll_100 <- c()
for(i in 1:5000){
  made <- n_hands(100, 1000, 100, 0.5)
  end_bankroll_100 <- c(end_bankroll_100, made[1])
}
mean(end_bankroll_100)
var(end_bankroll_100)
```
    e. the mean and variance of your bankroll after 500 hands (including busts).
```{r}
#In each simulation of 500 hands, your bankroll is added to a vector. Then the mean and variance is calculated.
end_bankroll_500 <- c()
for(i in 1:5000){
  made <- n_hands(500, 1000, 100, 0.5)
  end_bankroll_500 <- c(end_bankroll_500, made[1])
}
mean(end_bankroll_500)
var(end_bankroll_500)
```    

Note: you *must* stop playing if your player has gone bust. How will you handle this in the `for` loop?

2. **Markov Chains**. Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100.
    a. Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation.
```{r}
#This vector will contain the ending bankroll values after 100000 hands.
end_bankroll_100000 <- c()

p <- 0.48
#The 100000 hands are simulated 1000 times, after a winning hand the probability goes up by 0.01, after a losing hand the probability is reverted to 0.48.
for(j in 1:100){
bankroll = 0
for(i in 1:100000){
  if(p >= 1){
    p = 1
  }
    win <- rbinom(1,1,p)
    #The probability of winning is changed based on whether or not the current hand is won.
    if(win == 1){
      bankroll = bankroll + 100
      p = p + 0.01
    }else{
      bankroll = bankroll - 100
      p = 0.48
    }
}
  end_bankroll_100000[j] <- bankroll
}
end_bankroll_100000
#The mean values at the end are calculated.
mean(end_bankroll_100000)
#The mean is a relatively high negative value, so the game is rather unfair.
```
    b. Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically?
```{r}
#Same process, except now the starting probability is 0.5 rather than 0.48.
end_bankroll_100000 <- c()

p <- 0.50
for(j in 1:100){
bankroll = 0
for(i in 1:100000){
  if(p >= 1){
    p = 1
  }
    win <- rbinom(1,1,p)
    if(win == 1){
      bankroll = bankroll + 100
      p = p + 0.01
      
    }else{
      bankroll = bankroll - 100
      p = 0.50
    }
}
  end_bankroll_100000[j] <- bankroll
}

mean(end_bankroll_100000)
#The mean is a relatively high positive value, so the game again is unfair (but in your advantage!)

#Switch to starting probability of 0.49.
end_bankroll_100000 <- c()

p <- 0.49
for(j in 1:100){
bankroll = 0
for(i in 1:100000){
  if(p >= 1){
    p = 1
  }
    win <- rbinom(1,1,p)
    if(win == 1){
      bankroll = bankroll + 100
      p = p + 0.01
    }else{
      bankroll = bankroll - 100
      p = 0.49
    }
}
  end_bankroll_100000[j] <- bankroll
}

mean(end_bankroll_100000)
#This mean is much closer to 0. You could find a value for p that makes the game even more fair by looping through values from 0.480 to 0.500 and seeing which has the mean closest to 0.
```
    c. Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it. 
```{r}
#Same code as above, except for changing the increment to 2%.
p <- 0.48
for(j in 1:100){
bankroll = 0
for(i in 1:100000){
  if(p >= 1){
    p = 1
  }
    win <- rbinom(1,1,p)
    if(win == 1){
      bankroll = bankroll + 100
      #Increment is 2% now.
      p = p + 0.02
    }else{
      bankroll = bankroll - 100
      p = 0.48
    }
}
  end_bankroll_100000[j] <- bankroll
}

mean(end_bankroll_100000)
#Now the mean is quite large. Recall that when the increment was 1% the mean was negative. Testing a couple more values:

p <- 0.48
for(j in 1:100){
bankroll = 0
for(i in 1:100000){
  if(p >= 1){
    p = 1
  }
    win <- rbinom(1,1,p)
    if(win == 1){
      bankroll = bankroll + 100
      #Increment is 1.5% now.
      p = p + 0.015
    }else{
      bankroll = bankroll - 100
      p = 0.48
    }
}
  end_bankroll_100000[j] <- bankroll
}

mean(end_bankroll_100000)

#Now at 1.5% increment, the mean is still quite large. 

p <- 0.48
for(j in 1:100){
bankroll = 0
for(i in 1:100000){
  if(p >= 1){
    p = 1
  }
    win <- rbinom(1,1,p)
    if(win == 1){
      bankroll = bankroll + 100
      #Increment is 1.1% now.
      p = p + 0.011
    }else{
      bankroll = bankroll - 100
      p = 0.48
    }
}
  end_bankroll_100000[j] <- bankroll
}

mean(end_bankroll_100000)

#The increment being 1.1% gives a reasonably fair game.

```
