# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50.
```{r, eval = F}
#This loop prints the integers from 1 to 50. Let's not print all of them though,
for(i in 1:50){
  print(i)
}
```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
#This loop adds all the integers from 0 to 1000.
a <- 0
for(i in 1:1000){
  a <- a + i
}
a
```

B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
#This loop adds the even integers from 0 to 1000.
b <- 0
for(i in seq(0, 1000, by = 2)){
  b <- b + i
}
b
```

C. Now, repeat A and B WITHOUT using a loop.
```{r}
#The sum of the integers between 0 to 1000 without a loop.
sum(0:1000)
#The sum of the even integers between 0 to 1000 without a loop.
sum(seq(0, 1000, by = 2))
```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
#Begin by defining survey.clean to be equal to survey.
survey.clean <- survey
#The index i in the for loop refers to the columns 2 through 6, which are the answers by the 5 students.
for(i in 2:6){
  replacement <- survey[,i] #This column will replace the original in the dataframe.
  bad_answers <- setdiff(replacement, 1:5)
  indices <- which(replacement %in% bad_answers)
  replacement[indices] = NA #The unacceptable answers are found, their indices are determined and the replacement column puts NAs in place of the unaacceptable answers.
  survey.clean[, i] <- replacement #The original column is replaced in the cleaned dataframe.
}
survey.clean
```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
#The invalid.answers column is initialized with zeros.
survey.clean$invalid.answers <- 0
for(i in 1:6){
  #This piece of code adds the number of NAs in each row.
  survey.clean$invalid.answers[i] = sum(is.na(survey.clean[i,]))
}
survey.clean
```